---
title: "p8105_hw3_zw2974"
author: "Zihan Wu"
date: "2023-10-05"
output: github_document
---

```{r, include=FALSE}
library(tidyverse)
library(p8105.datasets)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom")) 
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem2
## Data cleaning:
* Format the data to use appropriate variable names;
* Focus on the “Overall Health” topic
* Include only responses from “Excellent” to “Poor”
* Organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
data("brfss_smart2010")
brfss = brfss_smart2010|> 
  janitor::clean_names() |> 
  filter(topic == "Overall Health"
         ) |> 
  mutate(response = factor(response, 
                           levels = c("Excellent", "Very good", "Good", "Fair", "Poor"),
                           order = T)
                           )
```

## Answering:
* In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
seven_02 =
  brfss |> 
  filter(
    year == 2002,
    response == "Excellent") |> 
  group_by(year, locationabbr) |> 
  summarize(n_county = n()) |> 
  filter(n_county >= 7) |> 
  pull(locationabbr)

seven_10 =
  brfss |> 
  filter(
    year == 2010,
    response == "Excellent") |> 
  group_by(year, locationabbr) |> 
  summarize(n_county = n()) |> 
  filter(n_county >= 7) |> 
  pull(locationabbr)
```
In 2002, states were observed at 7 or more locations: `r seven_02`

In 2010, states were observed at 7 or more locations: `r seven_10`

* Construct a dataset that is limited to `Excellent` responses, and contains, year, state, and a variable that averages the `data_value` across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the `geom_line` geometry and group aesthetic will help).

```{r}
excellent = 
  brfss |> 
  filter(response == "Excellent") |> 
  group_by(year, locationabbr) |> 
  summarize(avg = mean(data_value, na.rm = T))

excellent_plot =
  ggplot(
    excellent,
    aes(x = year, y = avg, color = locationabbr, group = locationabbr)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Spaghetti plot of  average value over time within states",
       x = "Year",
       y = "Average value of excellent responses")

excellent_plot
```
* Make a two-panel plot showing, for the years 2006, and 2010, distribution of `data_value` for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
ny_data = brfss %>%
  filter(locationabbr == "NY", year %in% c(2006, 2010))

ny_plot = ggplot(ny_data, aes(x=response, y=data_value)) +
  geom_boxplot() +
  facet_wrap(~year, scales = "free_x") +
  theme_minimal() +
  labs(title = "Distribution of health responses among locations in NY for 2006 and 2010",
       x = "Response",
       y = "Value")

print(ny_plot)

```

```{r}
ny_data = brfss %>%
  filter(locationabbr == "NY", year %in% c(2006, 2010))

ny_plot = ggplot(ny_data, aes(x=response, y=data_value)) +
  geom_boxplot() +
  facet_wrap(~year, scales = "free_x") +   # This line creates two panels
  theme_minimal() +
  labs(title = "Distribution of health responses among locations in NY for 2006 and 2010",
       x = "Response",
       y = "Value")

print(ny_plot)

```
